<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Full text}
-->

Full text
=====

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

Search functions in `rplos` can be used to get back full text in addition to any section of an article. However, if you prefer XML, this vignette is for you.

### Load package from CRAN

```{r eval=FALSE}
install.packages("rplos")
```

```{r}
library('rplos')
```

### Get full text URLs

Create urls for full text articles in PLOS journals

Here's the URL for XML full text for the DOI `10.1371/journal.pone.0086169`

```{r}
full_text_urls(doi='10.1371/journal.pone.0086169')
```

And for the DOI `10.1371/journal.pbio.1001845`

```{r}
full_text_urls(doi='10.1371/journal.pbio.1001845')
```

The function is vectorized, so you can pass in many DOIs

```{r}
full_text_urls(doi=c('10.1371/journal.pone.0086169','10.1371/journal.pbio.1001845'))
````

Use `searchplos()` to get a lot of DOIs, then get the URLs for full text XML

```{r}
dois <- searchplos(q = "*:*", fq='doc_type:full', limit=20)$data$id
full_text_urls(dois)
```

### Get XML

Get full text XML of PLOS papers given a DOI

```{r}
plos_fulltext(doi='10.1371/journal.pone.0086169')
```

`plos_fulltext()` is vectorized, so you can pass in more than one DOI

```{r}
plos_fulltext(c('10.1371/journal.pone.0086169','10.1371/journal.pbio.1001845'))
```

Get many DOIs, then index to get the full XML of the one you want (output not shown)

```{r}
dois <- searchplos(q = "*:*", fq='doc_type:full', limit=3)$data$id
out <- plos_fulltext(dois)
xml <- out[dois[1]][[1]]
```

Extract the abstract from the XML

```{r}
library("XML")
xpathSApply(xmlParse(xml), "//abstract/p", xmlValue)
```

Extract reference lists, just give back first one from each for brevity sake

```{r}
lapply(out[2:3], function(x){
 xpathApply(xmlParse(x), "//ref-list/ref")[[1]]
})
```

Make a text corpus

```{r}
library("tm")
out_parsed <- lapply(out, function(x){
 xpathApply(xmlParse(x), "//body", xmlValue)
})
tmcorpus <- Corpus(VectorSource(out_parsed))
(dtm <- DocumentTermMatrix(tmcorpus))
findFreqTerms(dtm, lowfreq = 50)
```
